Reproducible research - Course Project 1
=========================================

Format the notation of number inside text
```{r}
options(scipen=999, digits=2)
```

## Loading and preprocessing the data

1 - Download and load the data required for the assignment 

```{r}

link <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
temp <- tempfile()
download.file(link, temp)
unzip(zipfile = temp, exdir = getwd())

activity <- read.csv('activity.csv')

unlink(temp)
```

2 - Process the data???

## What is the mean total number of steps taken per day?

1 - Calculate the total number of steps taken per day

```{r}
stepsPerDay <- aggregate(activity$steps, by=list(date=activity$date), FUN = sum)
```

2 - Make a histogram of the total number of steps taken each day

```{r}
hist(stepsPerDay$x,
     breaks = 9,
     main = 'Number of steps taken per day',
     xlab = 'Steps per day')
```

3 - Calculate and report the mean and median of the total number of steps taken per day

```{r}
stepsTakenSummary <- summary(stepsPerDay$x)
stepsTakenMean <- unname(stepsTakenSummary[4])
stepsTakenMedian <- unname(stepsTakenSummary[3])
```

MEAN = `r stepsTakenMean`  
MEDIAN = `r stepsTakenMedian`  

## What is the average daily activity pattern?

1 - Make a time series plot of the 5-minutes interval

```{r}
stepsPerInterval <- aggregate(activity$steps, by=list(interval=activity$interval), FUN = mean, na.rm = TRUE)
with(stepsPerInterval,
     plot(interval, 
          x, 
          type='l', 
          ylab = 'Number of steps', 
          xlab = 'Interval', 
          main = 'Average number of steps taken in 5-minutes intervals'))
```

2 - Which 5-minutes interval contains the maximum number of steps? 

```{r}
maxStepsInterval <- stepsPerInterval[which(stepsPerInterval$x == max(stepsPerInterval$x)), 1]
```

The interval `r maxStepsInterval` contains the maximum number of steps

## Imputing missing values

1 - Calculate and report the number of missing values in the dataset

```{r}
totalNA <- sum(is.na(activity))
```

There are `r totalNA` rows with missing values  

2 - Strategy for filling the missing values   
  
In order to fill for the missing values, we will use the mean for that 5-minutes interval  

3 - Create a new dataset with NAs filled

```{r}
activityFilledNAs <- merge(activity, stepsPerInterval, by='interval')
activityFilledNAs$steps <- ifelse(is.na(activityFilledNAs$steps), 
                                  activityFilledNAs$x, 
                                  activityFilledNAs$steps)
```

4 - Make a histogram and report the mean and median

```{r}
stepsPerDayNoNAs <- aggregate(activityFilledNAs$steps, by=list(date=activity$date), FUN = sum)
hist(stepsPerDayNoNAs$x,
     breaks = 9,
     main = 'Number of steps taken per day',
     xlab = 'Steps per day')
stepsTakenSummaryNoNAs <- summary(stepsPerDayNoNAs$x)
stepsTakenMeanNoNAs <- unname(stepsTakenSummaryNoNAs[4])
stepsTakenMedianNoNAs <- unname(stepsTakenSummaryNoNAs[3])
```

MEAN = `r stepsTakenMeanNoNAs`  
MEDIAN = `r stepsTakenMedianNoNAs`  

The mean remains the same, however the median is lower when the NA values are filled

## Are there differences in activity patterns between weekdays and weekends?

1 - Create a new variable with two levels, weekend and weekday

```{r}
activity$week <- ifelse(weekdays(as.Date(activity$date)) == 'Saturday' | weekdays(as.Date(activity$date)) == 'Sunday', 'Weekend', 'Weekday')
```

2 - Make a panel plot

```{r}
stepsPerIntervalWE <- aggregate(activity$steps, by=list(interval=activity$interval, week=activity$week), FUN = mean, na.rm = TRUE)

with(stepsPerIntervalWE,
     plot(interval, 
          x, 
          type='l', 
          ylab = 'Number of steps', 
          xlab = 'Interval', 
          main = 'Average number of steps taken in 5-minutes intervals'))
```





